Interval-Valued Quantum Probabilities: Contextuality and the Born Rule

### 2017-05-16 prop_letter.tex
Updates from ShareLaTeX

Let $E_1$ and $E_2$ be to disjoint events with probabilities
$\mu(E_1)=[\ell_1,r_1]$ and $\mu(E_2)=[\ell_2,r_2]$. A first attempt
at calculating the probability of the combined event that $E_1$ or
$E_2$ occurs might be $\mu(E_1+E_2) = [\ell_1+\ell_2,r_1+r_2]$. In
some cases, this is indeed a sensible definition. For example, if
$\mu(E_1)=[0.1,0.2]$ and $\mu(E_2)=[0.3,0.4]$ we get
$\mu(E_1+E_2) = [0.4,0.6]$. But consider an event $E$ such that
$\mu(E)=[0.2,0.3]$ and hence $\mu(\overline{E})=[0.7,0.8]$. The two
events are disjoint; the previous na\"\i ve way of combining the
intervals would give $\mu(E+\overline(E))=[0.9,1.1]$ which is not a
valid probability interval. Moreover the event $E+\overline{E}$ is the
identity event covering the entire space; its probability interval
should be $[1,1]$ which is sharper than $[0.9,1.1]$. The problem is
that the two intervals are correlated: there is more information in
the combined event than in each event separately. In our example, even
though the true probability of $\mu(E)$ can be in anywhere in the
range $[0.2,0.3]$ and the true probability of $\mu(\overline{E})$ can
be in anywhere in the range $[0.7,0.8]$, the values are not
independent. Any value of $\mu(E) \leq 0.25$ will force
$\mu(\overline{E})\geq 0.75$. To account for such subtleties, the
axioms of interval-valued probability do not use a strict equality for
the combination of disjoint events. The correct constraint regarding
the probability of $E_1+E_2$ when $E_1$ and $E_2$ are disjoint is
$\mu(E_1+E_2) \subseteq [\ell_1+\ell_2,r_1+r_2]$ which, in combination
with the requirement for the identity event would force
$\mu(E+\overline{E})$ to always be $[1,1]$ irrespective of the
particular assignments of $\mu(E)$ or $\mu(\overline{E})$. When
combining non-disjoint events, there is a further subtlety whose
resolution will give the most general and final condition as
follows. For general, non-necessarily disjoint events $E_1$ and $E_2$,
we have:
\begin{equation}
\mu(E_1+E_2) + \mu(E_1E_2) \subseteq \mu(E_1) + \mu(E_2)
\label{eq:classicalconvex}
\end{equation}
which is a generalization of the classical inclusion-exclusion
principle that uses $\subseteq$ instead of $=$ for the same reason as
before. The new condition reduces to the previously motivated
condition when the events are disjoint, i.e., when $\mu(E_1E_2) = 0$.

>>>>

Let $E_1$ and $E_2$ be to disjoint events with probabilities
$\mu(E_1)=[\ell_1,r_1]$ and $\mu(E_2)=[\ell_2,r_2]$. A first attempt
at calculating the probability of the combined event that $E_1$ or
$E_2$ occurs might be $\mu(E_1+E_2) = [\ell_1+\ell_2,r_1+r_2]$. In
some cases, this is indeed a sensible definition. For example, if
$\mu(E_1)=[0.1,0.2]$ and $\mu(E_2)=[0.3,0.4]$ we get
$\mu(E_1+E_2) = [0.4,0.6]$. But consider an event $E$ such that
$\mu(E)=[0.2,0.3]$ and hence $\mu(\overline{E})=[0.7,0.8]$. The two
events are disjoint; the previous na\"\i ve way of combining the
intervals would give $\mu(E+\overline{E})=[0.9,1.1]$ which is not a
valid probability interval. Moreover the event $E+\overline{E}$ is the
identity event covering the entire space; its probability interval
should be $[1,1]$ which is sharper than $[0.9,1.1]$. The problem is
that the two intervals are correlated: there is more information in
the combined event than in each event separately. In our example, even
though the true probability of $\mu(E)$ can be in anywhere in the
range $[0.2,0.3]$ and the true probability of $\mu(\overline{E})$ can
be in anywhere in the range $[0.7,0.8]$, the values are not
independent. Any value of $\mu(E) \leq 0.25$ will force
$\mu(\overline{E})\geq 0.75$. To account for such subtleties, the
axioms of interval-valued probability do not use a strict equality for
the combination of disjoint events. The correct constraint regarding
the probability of $E_1+E_2$ when $E_1$ and $E_2$ are disjoint is
$\mu(E_1+E_2) \subseteq [\ell_1+\ell_2,r_1+r_2]$ which, in combination
with the requirement for the identity event would force
$\mu(E+\overline{E})$ to always be $[1,1]$ irrespective of the
particular assignments of $\mu(E)$ or $\mu(\overline{E})$. When
combining non-disjoint events, there is a further subtlety whose
resolution will give the most general and final condition as
follows. For general, non-necessarily disjoint events $E_1$ and $E_2$,
we have:
\begin{equation}
\mu(E_1+E_2) + \mu(E_1E_2) \subseteq \mu(E_1) + \mu(E_2)
\label{eq:classicalconvex}
\end{equation}
which is a generalization of the classical inclusion-exclusion
principle that uses $\subseteq$ instead of $=$ for the same reason as
before. The new condition reduces to the previously motivated
condition when the events are disjoint, i.e., when $\mu(E_1E_2) = 0$.
